---
name: Hayley
surname: Sowards
position: "Fellow at the National Cancer Institute, Department of Cancer Epidemiology and Genomics"
phone: +1 952 270 8803
www: hayleysowards.com
email: "hayley.sowards@gmail.com"
github: hsowards
linkedin: hayleysowards19
twitter: HayleyASowards
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tinytex)
library(knitr)
extrafont::loadfonts()
```




# Timeline Overview

```{r}
# All packages used below must be installed first
library(devtools)
library(lares)
library(ggplot2)

plot_timeline2 <- function(event, start, end = start + 1, label = NA, group = NA,
                           title = "Curriculum Vitae Timeline", subtitle = "Antoine Soetewey",
                           size = 7, colour = "orange", save = FALSE, subdir = NA) {
  df <- data.frame(
    Role = as.character(event), Place = as.character(label),
    Start = lubridate::date(start), End = lubridate::date(end),
    Type = group
  )
  cvlong <- data.frame(pos = rep(
    as.numeric(rownames(df)),
    2
  ), name = rep(as.character(df$Role), 2), type = rep(factor(df$Type,
    ordered = TRUE
  ), 2), where = rep(
    as.character(df$Place),
    2
  ), value = c(df$Start, df$End), label_pos = rep(df$Start +
    floor((df$End - df$Start) / 2), 2))
  maxdate <- max(df$End)
  p <- ggplot(cvlong, aes(
    x = value, y = reorder(name, -pos),
    label = where, group = pos
  )) + geom_vline(
    xintercept = maxdate,
    alpha = 0.8, linetype = "dotted"
  ) + labs( x = NULL, y = NULL, colour = NULL) +
    theme_lares2() + theme(panel.background = element_rect(
      fill = "white",
      colour = NA
    ), axis.ticks = element_blank(), panel.grid.major.x = element_line(
      size = 0.25,
      colour = "grey80"
    ))
  if (!is.na(cvlong$type)[1] | length(unique(cvlong$type)) >
    1) {
    p <- p + geom_line(aes(color = type), size = size) +
      facet_grid(type ~ ., scales = "free", space = "free") +
      guides(colour = FALSE) +
      scale_colour_hue()
  }
  else {
    p <- p + geom_line(size = size, colour = colour)
  }
  p <- p + geom_label(aes(x = label_pos),
    colour = "black",
    size = 2, alpha = 0.7
  )
  if (save) {
    file_name <- "cv_timeline.png"
    if (!is.na(subdir)) {
      dir.create(file.path(getwd(), subdir), recursive = T)
      file_name <- paste(subdir, file_name, sep = "/")
    }
    p <- p + ggsave(file_name, width = 8, height = 6)
    message(paste("Saved plot as", file_name))
  }
  return(p)
}

order <- c("Role", "Place", "Type", "Start", "End")
today <- as.character(Sys.Date())


### Edit from here ###
cv <- data.frame(rbind(
  c("BSc in Biology and Psychology", "ASU", "Academic", "2015-08-16", "2019-05-04"),
  c("NCI Fellow", "NIH", "Research Experience", "2019-08-05", today),
  c("Intern/Research Technician", "TGen", "Research Experience", "2018-06-10", "2019-07-26"),
  c("Research Assistant", "ATP", "Research Experience", "2017-03-01", "2019-05-04"),
  c("Research Assistant", "SPLAT", "Research Experience", "2016-08-16", "2018-08-01")
))
### Edit until here ###


colnames(cv) <- order
colour <- c("red", "blue")

plot_timeline2(
  event = cv$Role,
  start = cv$Start,
  end = cv$End,
  label = cv$Place,
  group = cv$Type,
  save = FALSE,
  subtitle = "Hayley Sowards" # replace with your name
)
```

\setlength{\parskip}{1em}

# Education
2019 - B.S. Biology (Genetics, Cellular, and Developmental) and Psychology 
- *Summa Cum Laude* 
- Arizona State University, Tempe, AZ 
  
2015 - Lakeville North High School <br>
	Lakeville, MN

# Research Experience
**Intramural Research Training Fellow at the NCI** August 2019 – Present
Functional analysis of melanoma risk loci
- Supervisors: Drs. Kevin Brown and Rebecca Hennessey
- **Independent Research:** Fine-mapping and functional analysis of multiple causal variants at the 9p21.3 melanoma risk locus

**Helios Scholar and Research Technician at TGen** June 2018 – July 2019

  Early life stress and effects on DNA methylation

  Supervisors: Drs. Matt Huentelman and Candace Lewis

  **Independent Research:** Associations Between Corporal Punishment and DNA Methylation in Stress Genes

**Research Assistant for the Arizona Twin Project** March 2017 – May 2019
	Childhood resilience and risk factors
	Supervisors: Drs. Kathryn Lemery-Chalfant, Mary Davis, and Leah Doane
	**Independent Research:** Dynamic Characteristics of Parent-Child Interactions: Mediating Role in Relation Between Parental Catastrophizing and Child Chronic Pain

**Research Assistant for the Shiota Psychophysiological Lab** August 2016 – August 2018
	Adult attachment, positive emotions, and ingroup relations
	Supervisor: Dr. Michelle Shiota

**Special Research Student at the Mayo Clinic** December 2017 – May 2018
	Identification of multiple myeloma candidate variants

# Publications
Lewis, C.R., Breitenstein, R. S., Henderson-Smith, A., Sowards, H. A., Piras, I. S., Huentelman, 
M., Doane, L.D. & Lemery-Chalfant, K. (2019) Dopaminergic Gene Methylation is 
Associated with Cognitive Performance in a Childhood Monozygotic Twin Study. 
Epigenetics, DOI: 10.1080/15592294.2019.1583032

Lewis, C.R., Sowards, H. A., Huentelman, M. & Lemery-Chalfant, K. (Under review) Early 
Parental Warmth Shapes General Health and Immune Epigenetics in Middle Childhood.

Lewis, C.R., Breitenstein, R. S., Henderson-Smith, A., Sowards, H. A., Huentelman, M., Doane, 
L.D. & Lemery-Chalfant, K. (In Preparation) Epigenetic Consequences of Harsh 
Parenting: NR3C1 and other HPA Gene Methylation Profiles in Middle Childhood.

Sowards, H. A., Clifford, S., Doane, L. D., Davis, M., & Lemery-Chalfant, K. (In Preparation) Dynamic Characteristics of Parent-Child Interactions: Mediating Role in Relation Between Parental Catastrophizing and Child Chronic Pain.

## Conference Abstracts
Lewis, C.R., Breitenstein, R. S., Henderson-Smith, A., Sowards, H. A., Huentelman, M., Doane, 
L.D. & Lemery-Chalfant, K. (2019). Epigenetic Consequences of Harsh 
Parenting: NR3C1 and other HPA Gene Methylation Profiles in Middle Childhood, Behavioral Genetics Association, Stockholm, Sweden, June 26th - 29th.

Lewis, C.R., Breitenstein, Sowards, H. A., Huentelman, M. & Lemery-Chalfant, K. (2019) 
Early Parental Warmth Shapes General Health and Immune Epigenetics in Middle 
Childhood, Society for Neuroscience, Chicago, IL, Oct 19th - 23rd.

## Poster Presentations
Sowards, H. A., Hennessey, R. C., Brown, K. B. (April 2020) Fine-mapping and functional 
analysis of multiple causal variants at the 9p21.3 melanoma risk locus. NIH 
Postbaccalaureate Poster Day, Bethesda, MD.

Sowards, H. A., Lewis, C. R., Breitenstein, R. S., Henderson-Smith, A., Doane, L.D., Lemery-
Chalfant, K., & Huentelman, M. (October 2018) Associations Between Corporal 
Punishment and DNA Methylation in Stress Genes. AZBio Awards, Phoenix, AZ.

Sowards, H. A., Lewis, C. R., Breitenstein, R. S., Henderson-Smith, A., Doane, L.D., Lemery-
Chalfant, K., & Huentelman, M. (July 2018) Associations Between Corporal Punishment 
and DNA Methylation in Stress Genes. Helios Scholars at TGen Research Symposium, Phoenix, AZ.

Lewis, C. R., Henderson-Smith, A., Styles, R., Sowards H. A., Doane, L., Lemery-Chalfant, K., 
Huentelman, M. (October 2018) KIBRA gene (WWC1) methylation is associated with 
cognitive performance in childhood.  Poster #1918, Presented at the 68th Annual 
Meeting of The American Society of Human Genetics (ASHG), San Diego, CA.

# Oral Presentations
“Associations Between Corporal Punishment and DNA Methylation in Stress Genes” (July 
2018; $250) Helios Scholars at TGen Research Symposium, Phoenix, AZ.

“Integrity in the Biosciences” (October 2018) Future Leader Presented by Helios at the 8th 
Integrity Summit, Phoenix, AZ. 

# Mentor Experience
Helios Scholars: Paulyn Kwak and Hannah Weisman	 	          June 2019 – July 2019
	Undergraduate Summer internship program at TGen

# Teaching Experience
Teaching Assistant: Physiological Psychology			    January 2019 – May 2019
	Supervisor: Dr. Eevin Jennings
Teaching Assistant: Psychology Writing Center	        	        August 2017 – December 2017
	Supervisor: Dr. Carolyn Cavanaugh-Toft

# Volunteer Experience
Maryland Science Olympiad: February 2020
North Star Tutoring: November 2019 – Present 
Task Manager for the Arizona Psychology Undergraduate Research Conference (AZPURC): October 2018 – February 2019
Sun Devil Safety Network Advisor: March 2016 – May 2019
Southwest Center for HIV/AIDS: March 2017 – May 2017

# Awards
Cancer Research Training Award: 2019 – 2021
College of Liberal Arts and Sciences Student Leader: 2018
ASU Alumni Association Legacy Scholarship: 2017 & 2018 
John W. Reich Prize in Applied Social Psychology: 2018
Dean’s Scholarship: 2015 – 2019
Dean’s List (> 3.5 GPA for semester): 2016 – 2019 

# References
**Dr. Kevin Brown**
Senior Investigator
National Cancer Institute
Division of Cancer Epidemiology & Genetics
Laboratory of Translational Genomics
8717 Grovemont Cir
Gaithersburg, MD 20887
Phone: (240) 760-6463
Email: brownkm2@nih.gov

**Dr. Mary Davis**
Professor and Associate Chair 
Arizona State University
Department of Psychology
PO Box 871104
Tempe, AZ 85287
Phone: (480) 965-2057
Email: mary.davis@asu.edu


**Dr. Matt Huentelman**
Professor
Translational Genomics Research Institute
Neurogenomics Division
445 N 5th St
Phoenix, AZ 85004
Phone: (602) 343-8730	
Email: mhuentelman@tgen.org
